---
title: "Imaging Impressions Overview"
author: "KH"
format: 
  html:
    code-fold: true
    df-print: paged
editor: visual
---

## MESSI Imaging Impressions Data Set

```{r}
#| echo: false
library(tidyverse)
library(discrim)
library(tidymodels)
imaging <- read.csv("SR8658598 - MESSI Study, IMPRESSION.csv")


```

```{r}
head(imaging)
```

### Date limitations

```{r}
imaging |> 
  mutate(ORDER_DATE = ymd_hms(ORDER_DATE)) |> 
  summarise(min_date = min(ORDER_DATE),
            max_date = max(ORDER_DATE))
```

### Regular expression/pattern matching imaging impressions

The string-searching method for defining positive/negative PE/DVT impression report

1.  **Search for 'key terms'**: pulmonary embol, thrombosis, dvt
2.  **Remove non-acute DVT thrombosis terms**: splenic vein, chronic thrombosis, portal venous, portal vein, basilar artery, transplant artery, cephalic vein, basilic vein, hepatic artery, superficial vein, hepatic vein
3.  **Prioritize specific negative impression reads and code as 'no clot'**: no evidence for an acute pulmonary embol, no evidence of clinically relevant acute pulmonary embol, not compatible with acute pulmonary embol
4.  **Second priority specific 'questionable' impression reads or impressions that I was unsure how to code**: questionable subsegmental pulmonary embol, questionable small right lower lobe subsegmental pulmonary embol, low quality examination for diagnosing pulmonary artery embol
5.  **Finally, code more generic terms positive impression reads as 'yes clot'**: +critical result: pulmonary embol, right upper lobe subsegmental pulmonary embol, acute pulmonary embol

```{r}
#| warning: false
time_zero <- readRDS("messi_demographic_time_zero_230815")
enroll_key <- readRDS("messi_enroll_key_230815")

names(imaging) <- tolower(names(imaging))

## Step 1: Grab IDs, CSN and time zero ----
step1 <- time_zero |> 
  select(messi_id,
         time_zero,
         hospital_admit_date,
         hospital_discharge_date)

step1_1 <- enroll_key |> 
  select(messi_id,
         mrn,
         csn)

step1_2 <- left_join(step1, step1_1, by = "messi_id") |> 
  distinct()

thrombosis <- inner_join(imaging, step1_2, by = "messi_id") |> 
  mutate(hospital_admit_date = ymd_hms(hospital_admit_date),
         hospital_discharge_date = ymd_hms(hospital_discharge_date),
         diff = as.numeric(difftime(order_date, hospital_admit_date, units = "days"))) |> 
  filter(diff >= 0 & diff <= 28 & order_date <= hospital_discharge_date) |> 
  ## Search for key terms 
  filter(grepl('pulmonary embol|thrombosis|dvt', impression, ignore.case = TRUE)) |> 
  ## Filter out common non-acute DVT thrombosis 
  filter(!grepl('splenic vein|chronic thrombosis|portal venous|portal vein|basilar artery|transplant artery|cephalic vein|basilic vein|hepatic artery|superficial vein|hepatic vein', impression, ignore.case = TRUE)) |> 
  ## Identify PE/DVT via impression strings 
  mutate(impression = tolower(impression),
         clot = case_when(grepl('no evidence for an acute pulmonary embol|no evidence of clinically relevant acute pulmonary embol|not compatible with acute pulmonary embol|no new acute pulmonary embol|no evidence of acute left upper extremity deep venous thrombosis|no definite new or large pulmonary embol|no evidence for an acute pulmonary embol|no central acute pulmonary embol|negative for acute pulmonary embol|chronic left lower lobe segmental pulmonary embol|no new acute pulmonary embol|no evidence for an acute pulmonary embol|no ct evidence of acute pulmonary embol|no new pulmonary embol|no new sites of pulmonary embol|previously seen pulmonary emboli in the left lower lobe has near resolved|no definite intraluminal pulmonary embol|no new pulmonary embol|no evidence of large central pulmonary embol|large central pulmonary embolus is not identified|no acute pulmonary, embol|no convincing evidence of pulmonary embol|no evidence for an acute pulmonary artery embol|low probability for acute pe|no large central or, new pulmonary embol|no acute pulmonary embol|no convincing evidence of acute pulmonary embol|no pulmonary embol|no acute central pulmonary embol|resolved pulmonary embol|no large pulmonary embol|no significant pulmonary embol|no definite pulmonary embol|no acute pulmonary embol|no definite evidence of pulmonary embol|no evidence of central or lobar pulmonary embol|no central or lobar pulmonary embol|no ct evidence of pulmonary embol|no acute pe|pulmonary embolism is excluded|no other pulmonary embol|no evidence of acute occlusive pulmonary embol|without large central pulmonary embol|no evidence of acute lower extremity deep venous thrombosis|no evidence of deep venous or superficial thrombosis|no evidence of acute rightlower extremity deep venous thrombosis|no evidence of acute left lower lower extremity deep venous thrombosis|no gross evidence of dvt|no evidence of acute vein thrombosis|no thrombosis|no gross pulmonary embol|no definite central pulmonary embol|no evidence of acute left  lower extremity deep venous thrombosis|no evidence of acute deep vein thrombosis|no evidence of deep vein thrombosis|no evidence of deep or superficial vein thrombosis|no evidence of venous thrombosis|no ct findings of pulmonary embol|dvt exam negative|no evidence of a large pulmonary embol|no evidence for pulmonary embol|no evidence of aortic dissection or central pulmonary embol|no evidence of acute proximal lle dvt|no evidence of acute right upper extremity deep venous thrombosis|no central, lobar or definite segmental pulmonary embol|no pulmonary embol|no left lower extremity proximal dvt|no evidence of acute right upper extremity deep venous thrombosis|no evidence of acute deep venous thrombosis|no evidence of acute bilateral lower extremity deep venous thrombosis|no evidence of central pulmonary embol|no evidence of vascular thrombosis|no findings of dural venous sinus thrombosis|no definite pulmonary embol|no large central pulmonary embolism seen|does not evaluate for dvt|no evidence of dural venous sinus thrombosis|no evidence of acute bilateral upper extremity deep venous thrombosis|without evidence of thrombosis|no deep venous thrombosis|no acute proximal segmental pulmonary embol|no ct angiographic evidence of pulmonary embol|negative for pulmonary embol|no evidence of acute right or left lower extremity deep venous thrombosis|no evidence of portal venous thrombosis|no evidence of dural venous sinus or deep cerebral venous thrombosi|no central pulmonary embol|no evidence of acute right lower extremity deep venous thrombosis|no pulmonary embol|no acute pulmonary embol|no gross pre-segmental pulmonary embol|no evidence of acute right lower extremity deep venous thrombosis|no evidence of acute pulmonary embol|no evidence of acute left lower extremity deep venous thrombosis|no evidence of pulmonary embol|no definite acute pulmonary embol|no central  pulmonary embol|no findings of acute pulmonary embol', impression) ~ 0,
                          grepl('questionable subsegmental pulmonary embol|questionable small right lower lobe subsegmental pulmonary embol|low quality examination for diagnosing pulmonary artery embol|nondiagnostic for pulmonary embol|suspicious for acute pulmonary embol|suspicious small pulmonary embol|suspicious for pulmonary embol|may represent pulmonary embol|limits evaluation for pulmonary embol|limited study for evaluation of pulmonary embol|suspicious small pulmonary embol|multiple small right sided pulmonary embol|if there is concern for pulmonary emboli|acute on chronic picture due , to pulmonary embolus|nondiagnostic study for pulmonary embol|recent pulmonary embol|infectious etiology such as septic pulmonary embol|setting of known pulmonary embol|history of right-sided pulmonary embol|setting of parenchymal disease or pulmonary embol|unchanged subacute pulmonary embol|indeterminate pulmonary emboli are not excluded|this study does not evaluate for pulmonary embol|essentially nondiagnostic assessment for peripheral pulmonary embol|previously demonstrated pulmonary embol|in the setting of pulmonary embol|chronic, nonocclusive deep vein thrombosis|detection of pulmonary embolism, limited by respiratory cardiopulmo|sequela of the pulmonary embol|changes of known extensive right-sided dvt|venous thrombosis is a consideration|question of cortical vein thrombosis|mca thrombosis|equivocal for isolated pulmonary embol|low quality examination for evaluation of pulmonary embol|pulmonary embolism versus artifact|known history of pulmonary embol|in the setting of known pulmonary embol|indeterminate pulmonary embol|suspicious for pulmonary embol', impression) ~ 1,
                          grepl('likely representing pulmonary embol|large proximal bilateral pulmonary embol|subsegmental right and possibly left lower lobe pulmonary embol|segmental left upper lobe pulmonary embol|subsegmental arterial pulmonary embol|nonocclusive multifocal pulmonary embol|subsegmental right upper lobe pulmonary embol|right-sided pulmonary embol|small pulmonary embol|interlobular pulmonary artery embol|new central left pulmonary embol|small, nonocclusive pulmonary embol|pulmonary embolus seen|lingular pulmonary embol|progression of pulmonary embol|multiple pulmonary embol|subsegmental pulmonary embol|pulmonary emboli are noted|acute nonocclusive pulmonary embol|acute/subacute new pulmonary embol|right upper lobe small pulmonary embol|representing multiple pulmonary embol|mild obstructive index pulmonary embol|left lower lobe proximal subsegmental pulmonary embol|acute central and peripheral pulmonary embol|saddle pulmonary embol|acute to subacute segmental left upper lobar pulmonary embol|acute, non-occlusive deep vein thrombosis|acute, occlusive deep vein thrombosis|acute,  non-occlusive deep vein thrombosis|acute,  occlusive deep vein thrombosis|suspected pulmonary embol|acute nonocclusive thrombosis of left axillary vein and one left brachial vein|right upper lobe segmental pulmonary embol|acute left upper extremity deep venous thrombosis|consistent with pulmonary embol|segmental pulmonary embol|favored to reflect deep venous thrombosis|right femoral vein deep venous thrombosis|acute subsegmental pulmonary embol|acute left-sided pulmonary embol|new acute left-sided pulmonary embol|acute bilateral pulmonary embol|in the setting of splenic vein thrombosis|new small subsegmental pulmonary embol|acute right-sided pulmonary embol|new posterior embol|nonocclusive deep venous thrombosis|unchanged portomesenteric vein thrombosis|extensive venous thrombosis|+critical result:  acute deep venous thrombosis|right middle lobe segmental pulmonary embol|right lower lobe pulmonary embol|ultrasound also demonstrating spontaneous thrombosis|thrombosis of the right hepatic vein|+critical result:  pulmonary embol|right upper lobe subsegmental pulmonary embol|multiple sites of nonocclusive thrombus|acute pulmonary embol|complete thrombosis of the ivc|partially occlusive deep venous thrombus|bilateral segmental acute pulmonary embol', impression) ~ 2)) |> 
  group_by(messi_id) |> 
  slice_max(clot) |> 
  mutate(clot_final = case_when(clot == 0 ~ 0,
                          clot == 1 ~ 8,
                          clot == 2 ~ 1),
         clot_final = as.factor(clot_final))

```

Total unique patients with pulmonary embol, thrombosis, dvt 'key words' within 28 days of hospital admission = 762

0 = negative impression report

1 = positive impression report

8 = questionable impression report

```{r}
summary(thrombosis$clot_final)
```

Benefits of this simple heuristic categorization compared to an NLP machine learning algorithm include easy to implement, maintain and audit.

### Compare to diagnosis codes

```{r}
#| echo: false
dx_codes <- readRDS("S:/Sepsis-ALI/Data Management/MESSI/data requests/Emilio Ramos/thrombosis/data/clotting_dx_v3")

dx_codes_limit <- inner_join(step1, dx_codes, by = "messi_id") |> 
  mutate(min = ymd_hms('2015-03-21 08:29:00')) |> 
  filter(hospital_admit_date >= min)

has_vte_codes <- dx_codes_limit |> 
  filter(vte == 1)
```

Limit the cohort to only those with with a hospital admit date after '2015-03-21 08:29:00' total = 2921

Have VTE based off of dx codes within time frame = 456 compared to imaging impressions = 106

```{r}
compare_clots <- full_join(has_vte_codes, thrombosis, by = "messi_id")

head(compare_clots)
```

## Pleural Effusion Classification Model

Using 1 month of CIRCE imaging impression data that I used a similar regular expression method to categorize the data into pleural effusion yes/no.

```{r}
library(tidyverse)
data <- readRDS("S:/Sepsis-ALI/Data Management/MESSI/xx-archive-projects/NLP imaging impressions/pleural_effusion_training_data")
```

```{r}
head(data)
```

## Pre-processing

Removes numbers and punctuation from free text

```{r}
data_preproc <- data |> 
  ## remove numbers
  mutate(impression = str_remove_all(impression, "[:digit:]"),
         ## remove punctuations
         impression = str_remove_all(impression, "[[:punct:]]"),
         effusion_yn = as.factor(effusion_yn))
```

```{r}
head(data_preproc$impression)
```

## Build First Classification Model

```{r}
library(tidymodels)

set.seed(1234)

data_split <- initial_split(data_preproc, strata = effusion_yn)

data_train <- training(data_split)
data_test <- testing(data_split)
```

```{r}
## dimensions of the two splits 
dim(data_train)
```

```{r}
dim(data_test)
```

```{r}
## additional pre-processing 
data_rec <-
  recipe(effusion_yn ~ impression, data = data_train)
```

'Tokenize' impressions into word tokens

```{r}
library(textrecipes)

data_rec <- data_rec %>%
  step_tokenize(impression) %>%
  ## only keep the 1000 most frequent tokens 
  step_tokenfilter(impression, max_tokens = 1e3) %>%
  step_tfidf(impression)
```

```{r}
data_wf <- workflow() %>%
  add_recipe(data_rec)
```

## Null model comparison

```{r}
set.seed(234)
data_folds <- vfold_cv(data_train)

data_folds
```

```{r}
null_classification <- null_model() %>%
  set_engine("parsnip") %>%
  set_mode("classification")

null_rs <- workflow() %>%
  add_recipe(data_rec) %>%
  add_model(null_classification) %>%
  fit_resamples(
    data_folds
  )
```

```{r}
null_rs %>%
  collect_metrics()
```

First shot of evaluation used a Naive Bayes model which was no better than the null model

## Evaluation of lasso classification model

"Least Absolute Shrinkage and Selection Operator"

Lasso regression adds a penalty term to a traditional logistic regression model so it can identify irrelevant or redundant variables

### Tuning lasso hyperparameters

Find the best penalty

```{r}
tune_spec <- logistic_reg(penalty = tune(), mixture = 1) %>%
  set_mode("classification") %>%
  set_engine("glmnet")

tune_spec
```

```{r}
lambda_grid <- grid_regular(penalty(), levels = 30)
lambda_grid
```

**V-fold cross validation**: randomly splits the data into V groups of roughly equal size

```{r}
tune_wf <- workflow() %>%
  add_recipe(data_rec) %>%
  add_model(tune_spec)

set.seed(2020)
tune_rs <- tune_grid(
  tune_wf,
  data_folds,
  grid = lambda_grid,
  control = control_resamples(save_pred = TRUE)
)

tune_rs
```

```{r}
collect_metrics(tune_rs)
```

```{r}
autoplot(tune_rs) +
  labs(
    title = "Lasso model performance across regularization penalties",
    subtitle = "Performance metrics can be used to identity the best penalty"
  )
```

```{r}
tune_rs %>%
  show_best("roc_auc")
```

Choose the model with the best ROC AUC within one standard error of the numerically best model

```{r}
chosen_auc <- tune_rs %>%
  select_by_one_std_err(metric = "roc_auc", -penalty)

chosen_auc
```

```{r}
final_lasso <- finalize_workflow(tune_wf, chosen_auc)

final_lasso
```

Run model on test data

```{r}
fitted_lasso <- fit(final_lasso, data_train)
```

```{r}
final_fitted <- last_fit(final_lasso, data_split)

collect_metrics(final_fitted)
```

```{r}
collect_predictions(final_fitted)  %>%
  roc_curve(truth = effusion_yn, .pred_TRUE, event_level = "second") %>%
  autoplot() +
  labs(
    color = NULL,
    title = "ROC curve for Positive PE Report",
    subtitle = "With final tuned lasso regularized classifier on the test set"
  )
```

```{r}
collect_predictions(final_fitted) %>%
  conf_mat(truth = effusion_yn, estimate = .pred_class) %>%
  autoplot(type = "heatmap")
```

```{r}
library(vip)

complaints_imp <- extract_fit_parsnip(final_fitted$.workflow[[1]]) %>%
  vip::vi(lambda = chosen_auc$penalty)

complaints_imp %>%
  mutate(
    Sign = case_when(Sign == "POS" ~ "Positive Imaging Report",
                     Sign == "NEG" ~ "Negative Imaging Report"),
    Importance = abs(Importance),
    Variable = str_remove_all(Variable, "tfidf_consumer_complaint_narrative_"),
    Variable = str_remove_all(Variable, "textfeature_narrative_copy_")
  ) %>%
  group_by(Sign) %>%
  top_n(20, Importance) %>%
  ungroup %>%
  ggplot(aes(x = Importance,
             y = fct_reorder(Variable, Importance),
             fill = Sign)) +
  geom_col(show.legend = FALSE) +
  scale_x_continuous(expand = c(0, 0)) +
  facet_wrap(~Sign, scales = "free") +
  labs(
    y = NULL,
    title = "Variable importance for predicting positive imaging reports for pleural effusion",
    subtitle = paste0("These features are the most important in predicting/n",
                      "whether an imaging report is positive or not for pleural effusion")
  )
```

**False negatives**: a good amount are initial classification errors on my part

```{r}
data_bind <- collect_predictions(final_fitted) %>%
  bind_cols(data_test %>% select(-effusion_yn))

data_bind %>%
  filter(effusion_yn == "TRUE", .pred_TRUE < 0.2) %>%
  select(impression) %>%
  slice_sample(n = 10)
```

False positives

```{r}
data_bind %>%
  filter(effusion_yn == "FALSE", .pred_FALSE < 0.2) %>%
  select(impression) %>%
  slice_sample(n = 10)
```

```{r}
## Save model to use for later 
saveRDS(fitted_lasso, pleural_effusion_fitted_model")
```
